---
title: "Introduction"
author: "Annelies Quinton"
date: '2022-09-14'
output: pdf_document
categories: []
tags: []
summary: Welcome to my weekly election blog for the upcoming 2022 midterms in November.
  The goal of this blog is to create a prediction for both the seat share and the
  popular vote for each party within the House of Representatives. I will use this
  blog to explain changes and additions to my prediction model. I will also provide
  graphics that support my claims and data reasoning.
slug: []
---



This first post serves as an introduction to this upcoming election by analyzing the outcomes of past recent midterms. Past elections can indicate patterns and trends that should be accounted for when predicting future elections. The data analyzed in this blog entry addresses voting outcomes at both the district and state level. Although districts elect Representatives, looking at statewide data can provide a more robust outlook at the trends of the state because factors such as redistricting and lack of opposition can limit the accuracy of district voting trends.

<br>

The first portion of this blog compares the vote share margin between the years of 2014 and 2020. As the shades of red increases, this represents the GOP having a larger percentage of the vote share in that state. Overall, between 2014 and 2020 we see a general decline in the GOP’s vote share as the color of the map decreases in brightness. In 2014, noteworthy states include Wyoming and Oklahoma which had the highest GOP vote share margin, 74.94% and 72.45%, respectively. Conversely, Massachusetts and Vermont had the lowest GOP vote share margin, 17.3% and 35.51%, respectively. For 2020, the two lowest states of the GOP vote share were in Massachusetts and Rhode Island, 21.97% and 26.01% respectively. The highest were in South Dakota, with 100% and Wyoming with 73.61%. South Dakota’s percentage can be explained by the fact that the one open House seat only had one candidate on the ballot, a republican, therefore eliminating the democratic vote share.  When comparing the mean GOP vote share across these years, we see that in 2014, the mean GOP vote share was 53.19%. Looking at 2020, this value dropped to 48.74%. These values corroborate the general visual trend of the map that shows the GOP vote share decreasing between the 2014 and 2020 elections. 


# Districts Vote Share

```{r}
# Hide code output
#knitr::opts_chunk$set(echo = FALSE)

# Loading libraries
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
library(tidyverse)
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
library(ggplot2)
install.packages("sf", repos = "http://cran.us.r-project.org")
library(sf)
install.packages("rmapshaper", repos = "http://cran.us.r-project.org")
library(rmapshaper)
install.packages("usmap", repos = "http://cran.us.r-project.org")
library(usmap)
states_map <- usmap::us_map()
install.packages("blogdown", repos = "http://cran.us.r-project.org")
library(blogdown)

# Reading in data

# Geographic data
get_congress_map <- function(cong=114) {
tmp_file <- tempfile()
tmp_dir <- tempdir()
zp <- sprintf("https://cdmaps.polisci.ucla.edu/shp/districts114.zip",cong)
download.file(zp, tmp_file)
unzip(zipfile = tmp_file, exdir = tmp_dir)
fpath <- paste(tmp_dir, sprintf("districtShapes/districts114.shp",cong), sep = "/")
st_read(fpath)
}

# 114th congress
cd114 <- get_congress_map(114)
cd114 <- cd114 %>% 
  mutate(DISTRICT = as.character(DISTRICT))%>%
  select(DISTRICT, STATENAME) %>% 
  rename(State = STATENAME)

# House data (1948 - 2020)
house <- read_csv(file = "house party vote share by district 1948 to 2020.csv")

# Joining House and CD114 data (2014) by district
# filter to 2014
h.2014 <- house %>% 
  filter(raceYear == 2014) %>% 
  select(raceYear, State, district_num, RepVotesMajorPercent, DemVotesMajorPercent) %>% 
  group_by(State, district_num) %>% 
  summarise(Rep_votes_pct = RepVotesMajorPercent) %>% 
  # rename district variable name to match shapefile
  rename(DISTRICT = district_num)

# change class
cd114$DISTRICT <- as.numeric(cd114$DISTRICT)

# Join election outcomes with shapefiles
cd114 <- cd114 %>% left_join(h.2014, by=c("DISTRICT", "State"))
cd114 <- ms_simplify(cd114, keep = 0.25, keep_shapes = T)
```


```{r}
# Plot districts
#district_map <- ggplot() +
 # geom_sf(data=cd114,aes(fill=Rep_votes_pct) ,
  #inherit.aes=FALSE,alpha=0.1) +
 # scale_fill_gradient(low = "white", high = "red", limits=c(0,100)) +
  #coord_sf(xlim = c(-172.27, -66.57), ylim = c(18.55, 71.23), expand = FALSE) +
  #theme_void()

#district_map
```

```{r}
# Voteshare margin by party (2020)
R_2020 <- house %>% 
  filter(raceYear == 2020) %>% 
  select(raceYear, State, district_num, district_id, RepVotes, DemVotes) %>% 
  # sumarize party vote by state
  group_by(State) %>% 
  # mutate Rep vote margin by state %>% 
  mutate(R_votemargin_st = (sum(RepVotes))/
           sum(RepVotes + DemVotes),
         D_votemargin_st = (sum(DemVotes))/
           sum(RepVotes + DemVotes)) %>% 
  rename(state = State)

#Voteshare margin by party (2014)
R_2014 <- house %>% 
  filter(raceYear == 2014) %>% 
  select(raceYear, State, district_num, district_id, RepVotes, DemVotes) %>% 
  # sumarize party vote by state
  group_by(State) %>% 
  # mutate Rep vote margin by state %>% 
  mutate(R_votemargin_st = (sum(RepVotes))/
           sum(RepVotes + DemVotes),
         D_votemargin_st = (sum(DemVotes))/
           sum(RepVotes + DemVotes)) %>% 
  rename(state = State)


# Plot voteshare margin 2020
 vs2020_map <- plot_usmap(data = R_2020, regions = "states", values = "R_votemargin_st", labels = TRUE) +
  scale_fill_gradient(low = "white", high = "red", name = "GOP two-party voteshare margin in 2020", limits=c(-0,1)) +
  theme_void()
 
 vs2020_map
 
# Plot voteshare margin 2014
vs2014_map <- plot_usmap(data = R_2014, regions = "states", values = "R_votemargin_st", labels = TRUE) +
  scale_fill_gradient(low = "white", high = "red", name = "GOP two-party voteshare margin in 2014", limits=c(0,1)) + 
  theme_void()

vs2014_map

# Gerrymandering

Seats_2020 <- read_csv("2020 - Sheet1.csv")
Seats_2020 <- Seats_2020 %>% 
  mutate(R_SeatShare = (RepWinner) / (RepWinner + DemWinner)) %>% 
  rename(state = Area)

gerry2020_map <- plot_usmap(data = Seats_2020, regions = "states", values = "R_SeatShare", labels = TRUE) +
  scale_fill_gradient(low = "white", high = "red", name = "GOP House Seat Share in 2020") + 
  theme_void()


gerry2020_map
```
```{r}
# Analysis
mean_vs_2014 <- mean(R_2014$R_votemargin_st)
mean_vs_2020 <- mean(R_2020$R_votemargin_st)

mean_ss_2020 <- mean(Seats_2020$DemWinner, na.rm=T) 
mean_ss_2020 <- (R_2020$R_votemargin_st)
```

